---
title: "HaloTag-MOR exploration"
output: html_notebook
---
#packages and utils
```{r}
library(tidyverse)
library(tidyr)
library(bio3d)
library(zoo)
library(patchwork)
library(pheatmap)


install.packages("pak")
library(pak)
pak::pak("jbengler/tidyheatmaps")
library(tidyheatmaps)


source("scripts/dms_analysis_utilities.R")
```

# import Enrich2 score files and annotate
```{r}

mor_scores_file = "data/DMS/main_identifiers_scores.tsv"

mor_variantscore_colnames <- c("hgvs", 
                               "SE_dms_1", "epsilon_dms_1", "score_dms_1",
                               "SE_dms_2", "epsilon_dms_2", "score_dms_2",
                               "SE_dms_3", "epsilon_dms_3", "score_dms_3",
                               "SE_dms_4", "epsilon_dms_4", "score_dms_4",
                               "SE_dms_5", "epsilon_dms_5", "score_dms_5",
                               "SE_dms_6", "epsilon_dms_6", "score_dms_6",
                               "SE_dms_7", "epsilon_dms_7", "score_dms_7",
                               "SE_dms_8", "epsilon_dms_8", "score_dms_8",
                               "SE_dms_9", "epsilon_dms_9", "score_dms_9",
                               "SE_dms_abundance", "epsilon_dms_abundance", "score_dms_abundance",
                               "SE_dms_internalization", "epsilon_dms_internalization", "score_dms_internalization",
                               "SE_dms_surface", "epsilon_dms_surface", "score_dms_surface")

mor_variantscore_coltypes <- c("hgvs", 
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric", "numeric")


mor_scores <- read_delim(mor_scores_file, 
                         col_names = mor_variantscore_colnames, 
                         skip=3)

mor_scores <- process_hgvs_df(mor_scores)

mor_scores$is.wt = mor_scores$mutation_type == "S"
```
# import sequence
```{r}

mor_wt = "MDSSAAPTNASNCTDALAYSSCSPAPSPGSWVNLSHLDGNLSDPCGPNRTDLGGRDSLCPPTGSPSMITAITIMALYSIVCVVGLFGNFLVMYVIVRYTKMKTATNIYIFNLALADALATSTLPFQSVNYLMGTWPFGTILCKIVISIDYYNMFTSIFTLCTMSVDRYIAVCHPVKALDFRTPRNAKIINVCNWILSSAIGLPVMFMATTKYRQGSIDCTLTFSHPTWYWENLLKICVFIFAFIMPVLIITVCYGLMILRLKSVRMLSGSKEKDRNLRRITRMVLVVVAVFIVCWTPIHIYVIIKALVTIPETTFQTVSWHFCIALGYTNSCLNPVLYAFLDENFKRCFREFCIPTSSNIEQQNSTRIRQNTRDHPSTANTVDRTNHQLENLEAETAPLP"
```

#import external data
```{r}
# import fortifications:

# import alpha missens .tsv file
mor_alpha_missense = read_tsv("data/alphamissense_OPRM_HUMAN.tsv")

# import csv of GPCRdb numbers
mor_gpcrdb = read.csv("data/GPCRdb_OPRM1_table.csv")

# import csv of conservation scores for non-olfactory class A
mor_conservation = read.csv("data/conservation_classA_nonolfactory_GPCRdb.csv")

# import gnomAD data csv
mor_gnomad = read.csv("data/gnomAD_OPRM1.csv")

# import atom distance to DAMGO (closest atom of ligand to atom of resi)
mor_damgo_distances = read.csv("data/damgo_ATOM_distances.csv")

# import G alpha distances (CA to CA)
mor_galpha_distances = read.csv("data/ga_CA_distances.csv")
```

#merge external data
# TODO: recalculate the Galpha distances to the MOR stucture and join to mor_scores, make the gnomad interpretable and join to mor_scores
```{r}
# join auxiliary data to the mor_scores datafame

# join the mor_gpcrdb to mor_scores by the pos column, keep all values in mor_scores
mor_scores = left_join(mor_scores, mor_gpcrdb, by = c("pos" = "pos"))

# join the mor_conservation to mor_scores by the GPCRdb column, keep all values in mor_scores
mor_scores = left_join(mor_scores, mor_conservation, by = c("GPCRdb" = "GPCRdb"))

# join the mor_alpha_missense to mor_scores by the hgvs column, keep all values in mor_scores
mor_scores = left_join(mor_scores, mor_alpha_missense, by = c("hgvs" = "hgvs"))

# join the mor_damgo_distances to mor_scores by the pos column, keep all values in mor_scores
mor_scores = left_join(mor_scores, mor_damgo_distances, by = c("pos" = "pos"))
```

#data annotations
```{r}
# add a new column to mor_scores called motif and fill every row with NA
# then fill rows corresponding to the following motifs and residue numbers
# create a new column called motifs in mor_scores and fill with NA
# DRY: 166-168
# NPxxY: 334-338
# PIF: 246, 157, 294
# CWxP: 294-297
# Na pocket: 116 
# disulfide: 142,219
# waters: 148, 223, 297
# TSS:356,357,358
# phospho: 365, 366, 372, 385, 396
# glyco: N40

mor_scores$motif = NA

# now fill motif column with motif name for pos above
mor_scores[mor_scores$pos %in% c(166, 167, 168), "motif"] = "DRY"
mor_scores[mor_scores$pos %in% c(334, 335, 336, 337, 338), "motif"] = "NPxxY"
mor_scores[mor_scores$pos %in% c(246, 157, 294), "motif"] = "PIF"
mor_scores[mor_scores$pos %in% c(294, 295, 296, 297), "motif"] = "CWxP"
mor_scores[mor_scores$pos %in% c(116), "motif"] = "Na+ pocket"
mor_scores[mor_scores$pos %in% c(142, 219), "motif"] = "disulfide"
mor_scores[mor_scores$pos %in% c(149, 223, 297), "motif"] = "waters"
mor_scores[mor_scores$pos %in% c(356, 357, 358), "motif"] = "TSS"
mor_scores[mor_scores$pos %in% c(377, 378, 379, 380, 381), "motif"] = "STANT"
mor_scores[mor_scores$pos %in% c(365, 366, 372, 385, 396), "motif"] = "phospho"
mor_scores[mor_scores$pos %in% c(40), "motif"] = "glyco"

# add columns for amino acid properties size, charge, etc
# https://www.imgt.org/IMGTeducation/Aide-memoire/_UK/aminoacids/IMGTclasses.html

# polarity
nonpolar<- c('F','W','I','L','M', 'V','C','A','G','P')
polar<- c('N','Q','S','H','Y','T','R','K','D','E')

#charge
charged<- c('K','R','H','D','E')
uncharged<- c('Y','S','T','N','Q')

#aromaticity
aromatic<- c('F','W','Y','H')
nonaromatic<- c('R','K','D','E','S','T','N','C','P','A','G','I','L','M','V')

# h-bonding
donor<- c('R','K','W')
acceptor<- c('D','E')
both<- c('N','Q','S','T','Y','H')
none<- c('A','C','G','I','L','M','F','P','V')

#hydropathy
hydrophobic<- c('F','W','I','L','M', 'V','C','A')
hydrophilic<- c('N','Q','D','E','K','R')
neutral<- c('Y','H', 'P', 'G', 'S','T')

#size
very_large<- c('F','W','Y')
large<- c('R','K','I','L','M')
medium<- c('V','H','Q','E')
small<- c('D','N','C','P','T')
very_small<- c('A','G','S')

#chemistry 
aliphatic<- c('V','A','L','I')
sulfur<- c('C','M')
hydroxyl<- c('S','T','Y')
basic<- c('K','R','H')
acidic<- c('D','E')
amide<- c('N','Q')


mor_scores <- mor_scores %>%
  mutate(aromaticity = ifelse(variants %in% aromatic, 'aromatic', 'nonaromatic'))
mor_scores <- mor_scores %>%
  mutate(polarity = ifelse(variants %in% polar, 'polar', 'nonpolar'))
mor_scores <- mor_scores %>%
  mutate(charge = ifelse(variants %in% charged, 'charged', 'uncharged'))
mor_scores <- mor_scores %>%
  mutate(h_bonding = case_when(
    variants %in% donor ~ 'donor',
    variants %in% acceptor ~ 'acceptor',
    variants %in% both ~ 'both',
    variants %in% none ~ 'none',
    TRUE ~ 'unknown'  # add this line to handle unexpected values
  ))
mor_scores <- mor_scores %>%
  mutate(hydropathy = case_when(
    variants %in% hydrophobic ~ 'hydrophobic',
    variants %in% hydrophilic ~ 'hydrophilic',
    variants %in% neutral ~ 'neutral',
    TRUE ~ 'unknown'  # add this line to handle unexpected values
  ))
mor_scores <- mor_scores %>%
  mutate(size = case_when(
    variants %in% very_large ~ 'hydrophobic',
    variants %in% large ~ 'hydrophilic',
    variants %in% medium ~ 'neutral',
    variants %in% small ~ 'small',
    variants %in% very_small ~ 'very_small',
    TRUE ~ 'unknown'  # add this line to handle unexpected values
  ))
mor_scores <- mor_scores %>%
  mutate(chemistry = case_when(
    variants %in% aliphatic ~ 'aliphatic',
    variants %in% sulfur ~ 'sulfur',
    variants %in% hydroxyl ~ 'hydroxyl',
    variants %in% basic ~ 'basic',
    variants %in% acidic ~ 'acidic',
    variants %in% amide ~ 'amide',
    TRUE ~ 'unknown'  # add this line to handle unexpected values
  ))

mor_scores <- mor_scores %>% filter(mutation_type != "X")

```

#import structures
# TODO: pull in non-DAMGO structures
```{r}
# pull in assorted pdb structures
damgo_8efq <- read.pdb("structures/8EFQ.pdb")
alphafold_active <- read.pdb("structures/alphafold_hmor_active.pdb")
alphafold_inactive <- read.pdb("structures/alphafold_hmor_inactive.pdb")
```

#heatmaps
```{r}
dms6_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_6,
             annotation_col = c(wt_aa, SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("blue","white","red"),
             color_na = "grey",
             color_legend_n = 1000,
             treeheight_row =50,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "dms6",
             filename = "output/plots/dms6_heatmap_annotated.pdf"
)

dms5_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_5,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("blue","white","red"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "dms5",
             filename = "output/plots/dms5_heatmap_annotated.pdf"
)

dms4_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_4,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("blue","white","red"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "dms4",
             filename = "output/plots/dms4_heatmap_annotated.pdf"
)

dms3_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_3,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("blue","white","red"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "dms3",
             filename = "output/plots/dms3_heatmap_annotated.pdf"
)

dms2_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_2,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("blue","white","red"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "dms2",
             filename = "output/plots/dms2_heatmap_annotated.pdf"
)

dms1_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_1,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("blue","white","red"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "dms1",
             filename = "output/plots/dms1_heatmap_annotated.pdf"
)

surface_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_surface,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("red","white","blue"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "surface",
             filename = "output/plots/surface_heatmap_annotated.pdf"
)

internalization_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_internalization,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             scale = "none",
             colors = c("red","white","blue"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             main = "internalization",
             filename = "output/plots/internalization_heatmap_annotated.pdf"
)

abundance_heatmap <- tidy_heatmap(mor_scores %>% filter(mutation_type %in% c("M", "S")) %>% filter(pos %in% 1:400) %>% arrange(pos), 
             rows = variants,
             columns = pos,
             values = score_dms_abundance,
             annotation_col = c(SSE, conservation, motif, damgo_distance),
             
             scale = "none",
             colors = c("red","white","blue"),
             color_na = "grey",
             color_legend_n = 1000,
             fontsize = 6,
             cellheight = 10,
             cellwidth = 10,
             height = 10,
             main = "abundance",
             filename = "output/plots/abundance_heatmap_annotated.pdf"
)

```

#density plots
```{r}
#plot densities of all scores

# Function to create density plots
create_density_plot <- function(data, score_column, label) {
  ggplot(data) +
    geom_density(aes(x = !!sym(score_column), color = mutation_type)) +
    xlab(label) +
    ylab("Density") +
    geom_vline(xintercept = 0, linetype = "dotted") +
    theme_classic() +
    labs(color = "Mutation type") +
    xlim(-3, 3) +
    ylim(0, 2)
}

# List to store density plots
density_plots <- list()

# Score columns and labels
score_columns <- c("score_dms_internalization", "score_dms_surface", "score_dms_abundance", 
                   "score_dms_1", "score_dms_2", "score_dms_3", "score_dms_4", "score_dms_5", "score_dms_6")

labels <- c("internalization", "surface", "abundance", "dms1", "dms2", "dms3", "dms4", "dms5", "dms6")

# Create density plots and store them in the list
for (i in seq_along(score_columns)) {
  density_plots[[i]] <- create_density_plot(mor_scores, score_columns[i], labels[i])
}

# Arrange plots using patchwork
combined_plots <- wrap_plots(density_plots, ncol = 3)

# Save the combined plot
ggsave("output/plots/master_density_plots.pdf", combined_plots, width = 10, height = 10)

```

#2d scatters with coloring
```{r}
# make scatter plots of all scores against all scores
plot_mor_scores <- function(data, x_var, y_var, color_var) {
  ggplot(data) +
    geom_point(aes(x = {{x_var}},
                   y = {{y_var}},
                   color = !!sym(color_var))) +
    theme_classic() + 
    labs(color = color_var) +
    xlim(-3, 3) +
    ylim(-3, 3)
}

# Define the list of variables
variables <- c("score_dms_surface", "score_dms_internalization", "score_dms_abundance", 
               "score_dms_1", "score_dms_2", "score_dms_3", 
               "score_dms_4", "score_dms_5", "score_dms_6")

# Define the list of color variables
color_variables <- c("motif", "conservation", "mutation_type", "alphamissense_classification", "SSE")

# Iterate over all combinations of x, y, and color variables
for (color_var in color_variables) {
  # Initialize an empty list to store scatter plots
  scatter_plots <- list()

  for (x_var in variables) {
    for (y_var in variables) {
      # Create the scatter plot and store it in the list
      scatter_plot <- plot_mor_scores(data = mor_scores, 
                                      x_var = !!sym(x_var), 
                                      y_var = !!sym(y_var),
                                      color_var = sym(color_var))
      scatter_plots <- c(scatter_plots, list(scatter_plot))
    }
  }

  # Combine the scatter plots using patchwork
  grid_plot <- wrap_plots(scatter_plots, ncol = 9)

  # Display the grid plot
  grid_plot

  # Save the grid plot as a PDF with a dynamic filename
  ggsave(paste0("output/plots/master_scatter_", color_var, ".pdf"), grid_plot, width = 40, height = 40)
}
```

#3d scatters
```{r}
# make a 3d scatter of the internalization, dms_3 and surface scores
install.packages("scatterplot3d")
library(scatterplot3d)
mor_3d <- scatterplot3d(mor_scores %>% filter(mutation_type != "X") %>% select(score_dms_internalization, score_dms_3, score_dms_surface), pch = 16, color = "blue", main = "3D Scatterplot of Internalization, DMS3 and Surface Scores")

install.packages("plotly")
library(plotly)
plot_ly(mor_scores %>% filter(mutation_type != "X"), x = ~score_dms_internalization, y = ~score_dms_3, z = ~score_dms_surface, color = ~motif, colors = "Set1", type = "scatter3d", mode = "markers", marker = list(size = 3))

# create a new df from mor_scores with the pos, each of the 9 dms scores, motif, mutation_type, and variants columns
mor_score_minimal <- mor_scores %>% select(pos, score_dms_1, score_dms_2, score_dms_3, score_dms_4, score_dms_5, score_dms_6, score_dms_surface, score_dms_internalization, score_dms_abundance, motif, mutation_type, variants, SSE)

# create a new df from mor_score_minimal and calculate the absolute sum of the scores 

```


